\begin{abstract}
LLM inference is increasingly bottlenecked by GPU memory:
    model sizes grow faster than device capacity,
    forcing each framework to implement its own complex, application-specific offloading.
CUDA Unified Virtual Memory (UVM) offers a transparent alternative,
    but is considered too slow for production---prior work reports
    over $5\times$ slowdowns relative to manual offloading.

We present {\uvmname}, a framework-aware UVM substrate for PyTorch
    that closes this gap without changing model code.
{\uvmname} contributes (1) a \emph{UVM-aware caching allocator}
    that pools and reuses UVM allocations to hide per-allocation driver overhead
    and returns cached memory to the driver under oversubscription
    to avoid the host OOM kills that a naive caching layer on UVM would suffer;
    and (2) \emph{framework-guided eviction},
    which marks partially-free cached blocks as reclaimable without copying data back to host.
For workloads where the active working set substantially exceeds GPU memory,
    we further study \emph{direct CPU-memory access} and
    \emph{access-counter--based migration} to mitigate page-fault thrashing.

We evaluate on Hugging~Face OPT and Qwen inference workloads
    across A100-PCIe, H100-PCIe, and GH200 NVLink-C2C.
When memory is not oversubscribed, {\uvmname} matches PyTorch's native
    \texttt{cudaMalloc}-based allocator within $20\%$.
Under oversubscription, where the native allocator crashes with OOM,
    {\uvmname} is $45\times$--$58\times$ faster than default UVM
    and $27\times$--$35\times$ faster than HF KV-cache offloading.
At high oversubscription, performance narrows to $1.6\times$--$1.7\times$ over default UVM.
\end{abstract}
