\section{Conclusion and Future Directions}
\label{chp:conclusion}

% \subsection{Summary}

This dissertation argues that the mismatch between modern hardware capabilities and
legacy system software designs is the root cause of critical bottlenecks in performance,
security, and extensibility---and that targeted, hardware- and workload-aware changes
to those software layers suffice to close the gap.

We identified three concrete manifestations of this mismatch.
First, commodity OS kernels remain hardwired to radix-tree page tables,
blocking experimentation with translation architectures suited to terabyte-scale servers.
Second, strict I/O memory protection in virtualized environments incurs 83--95\% throughput
degradation from per-DMA VM exits, forcing production deployments to disable it entirely.
Third, GPU memory oversubscription for AI workloads is unaddressed by the CUDA runtime,
leaving each framework to implement its own complex offloading logic or suffer
a more than 5$\times$ performance penalty from default UVM.

We address each problem with a targeted system.
\EMTbrand{} gives Linux an extensible, architecture-neutral interface for memory translation,
enabling OS support for radically different page table structures without modifying
the kernel for each new design.
Using \EMTbrand{}, we port ECPT and FPT---two recent hardware schemes for fast translation---to
Linux, reveal OS-level challenges invisible to hardware-only evaluation,
and optimize their designs from the OS perspective.
\vFreebrand{} redesigns the Linux IOMMU invalidation protocol around two mechanisms---one-for-many
invalidations and deferred page freeing---that amortize VM exit costs across concurrent unmap
requests while preserving strict security guarantees.
\vFreebrand{} achieves performance within 1--8\% of an unprotected system,
a 74--147$\times$ improvement over the state of the art.
\pytorchUVMbrand{} integrates a UVM-aware caching allocator and discard-based eviction
directly into PyTorch, enabling practical GPU memory oversubscription for LLM inference
without modifying application code.
Under oversubscription, \pytorchUVMbrand{} is 45$\times$--58$\times$ faster than default UVM
and 27$\times$--35$\times$ faster than hand-written KV-cache offloading.

Together, these systems validate the thesis:
bottlenecks in modern memory management arise not from fundamental hardware limitations
but from software abstractions designed without awareness of the hardware and workloads
they now serve.
Targeted changes to the system software layer---without hardware modification---are
sufficient to eliminate them.

\subsection{Unifying Lessons}

Across all three systems, a common structure emerges.
Each begins with a mature, widely deployed abstraction:
Linux's radix-tree paging, the Linux IOMMU subsystem, and CUDA Unified Virtual Memory.
Each abstraction is correct and general, but was designed under assumptions
that no longer hold at scale:
a single translation architecture, low per-operation overhead in virtualized environments,
and a driver-level view of workload access patterns.
The result in each case is a semantic gap between what the abstraction exposes
and what hardware and workloads actually need.

The key insight is that closing this gap does not require wholesale replacement
of the existing stack.
\EMTbrand{} retains Linux's page fault handling, huge page management, and
user-space memory interfaces while adding a thin extensibility layer beneath them.
\vFreebrand{} retains the IOMMU hardware protocol while reorganizing the software
path that serializes invalidation requests.
\pytorchUVMbrand{} retains CUDA UVM's transparency guarantee while adding
allocator-level and eviction-level awareness of framework caching behavior.
In each case, the fix is surgical: identify the specific invariant that the
legacy design enforces unnecessarily, relax it in a principled way,
and recover the lost performance or extensibility.

A second lesson is the importance of cross-layer evaluation.
Hardware architecture proposals for new translation schemes rely on simulation
models that assume OS overhead is constant across designs---an assumption \EMTbrand{}
shows to be false.
IOMMU hardware supports strict invalidation, but its cost is determined entirely
by how the hypervisor software serializes the resulting VM exits,
a detail \vFreebrand{} exposes.
UVM page fault latency is dominated not by hardware migration cost but by the
interaction between the driver's eviction policy and the framework's caching allocator,
the insight that motivates \pytorchUVMbrand{}.
Effective system design requires evaluating the full hardware--software stack,
not just the hardware layer in isolation.

% \subsection{Future Directions}

% The three systems in this dissertation open several directions for future work.

% \subsubsection{Extensible Translation for Heterogeneous Devices}
% \EMTbrand{} focuses on CPU memory translation, but the same extensibility problem
% exists for other translation-capable devices:
% IOMMUs, NPU memory management units, and near-memory processing units
% each use distinct page table organizations.
% Extending \EMTbrand{}'s architecture-neutral interface to cover non-CPU devices
% would allow a unified OS framework to manage memory across the full heterogeneous stack.
% The increasing diversity of interconnected memory devices---CXL-attached memory,
% disaggregated memory pools---makes OS extensibility for translation
% an increasingly pressing requirement.

% \subsubsection{Application-Aware Eviction for UVM}
% As demonstrated in our evaluation, thrashing remains a primary bottleneck
% for \pytorchUVMbrand{} under heavy oversubscription.
% The root cause is a semantic gap between the UVM driver's eviction policy and
% the framework's caching allocator: the driver evicts pages based on hardware-level
% access history, which frequently conflicts with logical reuse patterns visible only
% at the framework level.
% Closing this gap requires eviction policies informed by application-level data lifetimes.
% Existing offloading frameworks already compute detailed schedules for when data
% will be needed next; passing these schedules down as hints to the UVM driver
% would allow the driver to prefetch and evict more effectively
% while retaining UVM's transparency guarantee.

% \subsubsection{Multi-GPU and Cross-Node UVM}
% \pytorchUVMbrand{} is currently validated on single-GPU configurations.
% Production AI workloads require multi-GPU and cross-node deployments,
% where UVM semantics become more complex:
% every GPU can access a shared managed address space,
% and each GPU's caching allocator must reason about data resident on other devices.
% Extending \pytorchUVMbrand{} to multi-GPU systems requires coordinating
% eviction decisions across allocators and understanding how NVLink and CXL
% interconnect topologies affect migration cost.

% \subsubsection{Hardware--Software Co-Design for IO Memory Protection}
% \vFreebrand{} achieves near-native performance for strict vIOMMU protection
% through software reorganization alone.
% Future hardware could reduce costs further by exposing explicit batching
% interfaces for IOTLB invalidation, eliminating the need for the one-for-many
% aggregation that \vFreebrand{} implements in software.
% Similarly, hardware support for deferred page table reclaim---analogous
% to RCU in the CPU memory subsystem---could reduce the dedicated-core overhead
% that \vFreebrand{} currently requires.
% These directions represent opportunities for hardware--software co-design
% guided by the bottlenecks identified in this work.

\subsection{Closing Remarks}

Modern computing systems are hardware-rich and software-constrained.
The hardware advances of the last decade---terabyte DRAM, hundred-gigabit
device interconnects, and high-bandwidth GPU memory---are available in today's
servers, but legacy system software layers prevent workloads from exploiting them fully.
This dissertation demonstrates that the constraint is not fundamental:
the bottleneck is the abstraction layer, not the hardware,
and targeted changes to system software suffice to close the gap.

We hope the systems presented in this dissertation---\EMTbrand{}, \vFreebrand{},
and \pytorchUVMbrand{}---serve both as practical tools and as evidence for a
broader research agenda: that hardware- and workload-aware system software redesign
is a high-leverage approach to unlocking the full potential of modern memory hardware.
