\section{Introduction}
\label{chp:intro}

% \subsection{The Modern Computing Stack: Rich Hardware, Aging Foundations}

Modern computing bears little resemblance to that of a decade ago.
A modern datacenter is no longer a warehouse of identical servers running batch
jobs; it is a heterogeneous fabric of CPUs, GPUs, accelerators, smart NICs, and
disaggregated memory pools, knitted together by hundred-gigabit interconnects
and shared by millions of tenants.
Server DRAM has grown from gigabytes to terabytes per socket;
GPUs, once graphics coprocessors, have become the dominant substrate for AI,
linked by cache-coherent interconnects such as NVLink-C2C\@;
and DMA-capable devices---NICs, NVMe drives, accelerators---now move data at
hundreds of gigabits per second without ever involving the CPU\@.

In parallel with the hardware evolution,
workloads have also grown in scale and complexity.
In memory databases,
can easily generate a memory working set of hundreds of terabytes~\cite{redis},
putting greater pressure on memory management and address translation.
Network-intensive applications in virtualized clouds issue DMA at rates that
saturate a DRAM channel of the host~\cite{connectx7},
putting relentless pressure on the OS IOMMU subsystem that mediates every device access.
A single LLM training run marshals tens of thousands of GPUs across hundreds of
hosts to fit a model whose parameter count has grown by orders of magnitude in
five years~\cite{llama3}; 

Across this hardware and workload landscape, memory is the central resource that ties the
stack together.
Every workload, on every device, ultimately reads and writes memory:
a CPU thread chases pointers through a terabyte heap;
a NIC streams packets directly into guest physical memory;
a GPU kernels perform tensor computations by reading from and writing to device and host memory.
Every hardware advance---in processors, accelerators, interconnects, or memory
technology itself---must eventually be reflected in how the software stack
\emph{maps}, \emph{protects}, and \emph{moves} memory on behalf of those workloads.
Memory is the join point between hardware capability and application performance.

Yet, the \emph{system software} layers that manage this memory---chiefly the
operating system kernel, together with the hypervisor, device drivers, and
AI frameworks that sit alongside it---have not kept pace.
The kernel's memory management subsystem, IOMMU subsystem, and the
device memory management system were each conceived for a world that no longer exists:
servers with limited memory capacity,
device interconnects orders of magnitude slower than today's,
and accelerators with dedicated memory large enough to hold the entire working set.
Their core abstractions---the radix-tree page table, the synchronous IOTLB invalidation,
the application-opaque managed-memory pool---were correct and general for the hardware of their day.
At today's scale and speed, they have become the bottleneck.

\subsection{The System Software Gap}

This dissertation argues that the most consequential bottlenecks in modern memory
systems are \emph{system software}:
specifically, of OS and runtime abstractions designed without the hardware
or workloads they now face.
% , and not extensible enough to absorb the mismatch
% gracefully.
The cost of the mismatch is paid in three currencies:
\textbf{performance}, when software overhead consumes the headroom that new
hardware provides;
\textbf{security}, when production systems are forced to disable expensive
protection mechanisms to remain usable;
and \textbf{extensibility}, when hardwired software designs prevent the deployment
of new hardware ideas, slowing the broader research and development cycle.
Across the stack---from CPU virtual memory, to I/O memory protection, to GPU
device memory---the same pattern recurs:
a mature, widely deployed abstraction that was once correct and general now
exposes a semantic gap between what the hardware can do and what the workload
needs.

This dissertation studies that gap in three concrete settings, each chosen
because it is both representative of the broader pattern and important in its
own right:
CPU virtual memory translation in the OS kernel,
I/O memory protection in the OS kernel (specifically, the guest kernel's IOMMU
subsystem, in concert with the host kernel and hypervisor),
and the GPU memory management system.
In each setting, we identify the specific invariant that the legacy abstraction
enforces unnecessarily, redesign the software layer around that finding, and
show that targeted, hardware- and workload-aware system-software changes are
sufficient to close the gap---without modifying the hardware, refactoring the applications, or
abandoning the existing system stacks.
% \subsection{Three Manifestations of the Gap}

% The bottleneck manifests in three distinct but related problems,
% spanning CPU memory, I/O memory, and device memory:
\begin{enumerate}
\item \textbf{CPU virtual memory translation} has become a major bottleneck for
    memory-intensive workloads on terabyte-scale servers.
    Address translation can consume 20--50\% of execution time on big-memory
    workloads~\cite{basu2013efficient,gupta2021midgard,kwon2016coordinated,margaritov2019asap},
    and a substantial body of computer-architecture research has
    proposed faster translation schemes---hash-based~\cite{skarlatos2020ecpt,yaniv2016hash,jovan2022nestedecpt},
    tree-based~\cite{park2022fpt}, or hybrid~\cite{alam2017doityourself,kanellopoulos2023utopia}---that
    promise to recover much of this overhead.
    Yet commodity OS kernels remain hardwired to a single translation
    architecture, the radix-tree page table, and provide no extensible interface
    to experiment with alternative designs.
    The difficulty of building OS support has, in turn, discouraged disruptive
    architecture research: hardware proposals are evaluated almost entirely in
    simulation, under the assumption that OS overhead is constant across
    translation architectures---an assumption we show to be false.

\item \textbf{I/O memory protection in virtualized environments} is prohibitively
    expensive.
    In modern clouds, virtual machines are routinely granted direct access to
    high-performance devices, but allowing a guest-controlled device direct
    access to physical memory poses severe intra-guest security risks.
    Existing IOMMU mechanisms offer a strong, ``strict,'' safety guarantee that
    purges stale device translations immediately after each DMA;
    however, in virtualized environments, enforcing this guarantee requires at
    least one VM exit per DMA to invalidate translation entries in IOMMU caches.
    The resulting overhead is staggering: 83--95\% application throughput
    degradation---a penalty so severe that production deployments universally
    disable strict protection, leaving them vulnerable to known attack classes.
    Cloud operators are thus forced into a choice between performance and
    security that the hardware itself does not require.

\item \textbf{GPU memory oversubscription for memory-intensive AI workloads}
    has become an existential issue for the AI software stack.
    Flagship model sizes grow faster than GPU memory capacity, and the dominant
    response has been for each AI framework to implement its own complex,
    application-specific offloading logic---hand-written code paths for
    KV-cache eviction, parameter swapping, and activation rematerialization,
    re-implemented across PyTorch~\cite{pytorch}, vLLM~\cite{vllm}, DeepSpeed~\cite{deepspeed}.
    The CUDA runtime offers a transparent, driver-level alternative---Unified
    Virtual Memory (UVM)---but it is widely considered impractical for
    production.
    Our measurements show that UVM is over 50$\times$ slower than PyTorch's
    native \texttt{cudaMalloc}-based allocator when memory is not oversubscribed,
    and over 5$\times$ slower than application-level offloading under
    oversubscription (consistent with prior work~\cite{infinigen}).

    % prior work reports it is more than 5$\times$ slower than
    % application-level offloading~\cite{infinigen}, because the runtime has no visibility into
    % high-level access patterns and the framework's caching allocator actively
    % fights it.
\end{enumerate}

These three problems span very different parts of the stack---a kernel data
structure, a kernel-level invalidation protocol that spans guest and host,
and a GPU memory management system---but they share a common shape.
In each, modern hardware exposes a capability that the workload demands;
in each, a legacy software abstraction prevents the workload from using it;
and in each, the cost of the mismatch is paid in performance, security, or
the engineering effort required to work around the abstraction.

\subsection{Thesis Statement}

\noindent\emph{Modern memory systems require transformation to meet the demands
of contemporary hardware and workloads; making the system software layers aware
of both---through targeted, surgical changes---eliminates the bottlenecks of
performance, security, and extensibility that arise from legacy designs.}

\subsection{Approach: Three Systems}

This dissertation supports the thesis with three systems, each addressing one of
the manifestations described above.
Each system retains the surrounding stack---the OS kernel, the virtualization
stack it participates in, or the GPU memory management system---but adds a thin
layer of awareness that closes the gap between the abstraction and the underlying
hardware or workload.

\paragraph{EMT: An OS Framework for New Memory Translation Architectures.}
The first problem---hardwired radix-tree translation in Linux---blocks
researchers and hardware designers from experimenting with new translation
architectures.
New schemes such as hash-table-based ECPT~\cite{skarlatos2020ecpt} and
flat FPT~\cite{park2022fpt} have been shown to reduce translation overhead
significantly in simulation, but almost no proposals have ever run on a commodity OS.
Without OS experiments, hardware research relies on performance models that
assume OS overhead is constant across translation architectures---an assumption
we show to be false.

We present \emph{Extensible Memory Translation (EMT)}, a pragmatic framework
atop Linux to empower different hardware schemes of memory translation
such as radix tree and hash table.
EMT provides an architecture-neutral interface that
1) supports diverse memory translation architectures,
2) enables hardware-specific optimizations,
3) accommodates modern hardware and OS complexity,
and 4) has negligible overhead over hardwired implementations.
We port Linux's memory management onto EMT and show that
EMT enables extensibility without sacrificing performance.
We use EMT to implement OS support for ECPT and FPT, two recent experimental
translation schemes for fast translation;
EMT enables us to understand the OS perspective of these architectures
and further optimize their designs.

\paragraph{Virtually Free: Making Strict IO Memory Protection Feasible for Virtualized Environments.}
The second problem concerns IO memory protection in virtualized environments.
Device pass-through delivers near-native I/O performance to virtual machines,
but allowing a guest-controlled device direct access to physical memory poses
severe intra-guest security risks.
Virtualized IOMMUs (vIOMMUs) address this by exposing a virtual IOMMU to the
guest, enabling fine-grained DMA protection.
However, enforcing the highest level of protection---\emph{strict} invalidation,
which purges IOTLB entries immediately after each DMA---requires
a synchronous VM exit per unmap in current hardware-assisted designs.
A VM exit costs 10--20\,$\mu$s; a bare-metal IOTLB invalidation costs 200--300\,ns.
The result is 83--95\% application throughput degradation---a penalty so severe
that production deployments universally disable strict protection.

We present {\vFreebrand} a simple modification to existing IO memory protection
mechanisms that makes strict IO memory protection feasible in virtualized environments.
The key insight in {\vFreebrand} is that, in virtualized environments,
it is possible to reduce the overheads of strict IO memory protection
to a bare minimum by \textit{increasing} the cost of address translation.
Intuitively, this is because each VM exit has much higher cost
than the worst-case translation cost.
{\vFreebrand} design incorporates two ideas---\textit{one-for-many invalidations}
and \textit{deferred free page}---that, when put together, result in higher
address translation cost but significantly reduce the number of VM exits
(and hence, IO memory protection overheads) while maintaining strict IO memory protection.
We implement {\vFreebrand} on Linux, and evaluate it across multiple real-world applications.
Our evaluation suggests that, at the cost of one dedicated core,
{\vFreebrand} consistently achieves performance within 1--8$\%$ of
the unprotected system---74--147$\times$ higher than state-of-the-art IO memory protection mechanisms.

\paragraph{PyTorch-UVM: Practical Unified Virtual Memory for Large Language Model Inference.}
The third problem is GPU memory management for LLM inference.
Flagship model sizes grow faster than GPU memory capacity,
yet the transparent alternative provided by the CUDA driver---Unified Virtual
Memory (UVM)---has been considered impractical for production: prior work
reports UVM is more than 5$\times$ slower than application-level offloading.
The gap has two sources.
First, ML frameworks like PyTorch bypass UVM with their own caching allocators,
so managed allocations miss pooling and reuse optimizations.
Second, the driver has no visibility into model execution order,
so page faults and migrations are triggered late and often evict data that is
about to be reused.

We present {\pytorchUVMbrand}, an application-aware UVM substrate for PyTorch
that closes this gap without changing model code.
{\pytorchUVMbrand} contributes (1) a \emph{UVM-aware caching allocator}
that proactively returns cached memory to the driver under oversubscription,
avoiding the OOM kills that naive UVM caching suffers,
and (2) \emph{discard-based eviction},
which marks cached blocks as reclaimable without copying data back to host.
We evaluate on Hugging~Face OPT and Qwen inference workloads
across A100-PCIe, H100-PCIe, and GH200 NVLink-C2C.
When memory is not oversubscribed, {\pytorchUVMbrand} matches PyTorch's native
\texttt{cudaMalloc}-based allocator within $20\%$.
Under oversubscription, where the native allocator crashes with OOM,
{\pytorchUVMbrand} is $45\times$--$58\times$ faster than default UVM
and $27\times$--$35\times$ faster than HF KV-cache offloading.
At high oversubscription, performance narrows to $1.6\times$--$1.7\times$ over default UVM.

\subsection{Summary of Contributions}

This dissertation makes the following contributions:

\begin{itemize}
    \item \textbf{EMT} (\emph{EMT: An OS Framework for New Memory Translation Architectures}):
        A framework that gives Linux an extensible, architecture-neutral
        interface for memory translation, enabling OS support for radically different
        page table structures without forking the kernel; plus an open platform for
        evaluating OS kernels on new translation hardware.

    \item \textbf{Virtually Free} (\emph{Virtually Free: Making Strict IO Memory Protection
        Feasible for Virtualized Environments}):
        A redesign of the Linux IOMMU subsystem that
        makes strict I/O memory protection feasible in virtualized environments for
        the first time without sacrificing throughput.

    \item \textbf{PyTorch-UVM} (\emph{PyTorch-UVM: Practical Unified Virtual Memory for
        Large Language Model Inference}):
        An application-aware UVM integration for PyTorch that
        enables practical GPU memory oversubscription for LLM inference without changing application code.

    \item \textbf{A unifying perspective.} Across all three systems, the bottleneck is
        not the hardware itself but the software systems layer that sits between
        hardware capabilities and application demands.
        Together, these systems demonstrate that making system software layers hardware-
        and workload-aware---through targeted changes---eliminates bottlenecks in
        performance, security, and extensibility.
\end{itemize}

\subsection{Dissertation Roadmap}

The remainder of this dissertation is organized as follows.
\ref{chp:emt} presents EMT and its evaluation on new translation architectures.
\ref{chp:viommu} presents the redesigned vIOMMU mechanism and its evaluation
on 400\,Gbps networking workloads.
\ref{chp:uvmtorch} presents Pytorch-UVM and its evaluation on LLM inference across
PCIe-based and NVLink-C2C hardware.
\ref{chp:conclusion} concludes and discusses future directions.
